---
title: "Xu_Week_07"
author: "Yankun Xu"
date: "November 5th, 2016"
output: html_document
---
Let's load package first
```{r}
library(tidyverse)
library(readxl)
library(stringr)
library(choroplethr)
library(dplyr)
library(stringr)
library(ggplot2)
setwd("C:/Users/xuyankun/Xu_Week_07")
```

- Download the csv file on maternal mortality at
    http://apps.who.int/gho/athena/data/xmart.csv?target=GHO/MDG_0000000025,MDG_0000000026&profile=crosstable&filter=COUNTRY:*;YEAR:*&x-sideaxis=COUNTRY;YEAR&x-topaxis=GHO
and store it in a data directory in your project repo. (You can read about the data  at http://apps.who.int/gho/data/view.main.SDG31v?lang=en )
- Although the data is tidy, it needs cleaning. Notice that spaces are used instead of commas in numbers. Turn the numbers for maternal mortality inside square brackets into two new variables, lower_bound and upper_bound

```{r}
data1<-read.csv("xmart.csv")
data1_tidy<-data1
colnames(data1_tidy)[3]<-"Maternal.Mortality"
colnames(data1_tidy)[4]<-"Sklled.Health"

data1_tidy$Maternal.Mortality<-str_replace_all(data1_tidy$Maternal.Mortality, " ","")

#get lower bound
data1_tidy$Lower_bound<-str_extract_all(data1_tidy$Maternal.Mortality, "\\[[0-9]+")
data1_tidy$Lower_bound <- str_replace(data1_tidy$Lower_bound , "\\[" , "")
data1_tidy$Lower_bound<- as.numeric(data1_tidy$Lower_bound)

#get upper bound
data1_tidy$Upper_bound<-str_extract_all(data1_tidy$Maternal.Mortality, "[0-9]+\\]")
data1_tidy$Upper_bound <- str_replace(data1_tidy$Upper_bound , "\\]" , "")
data1_tidy$Upper_bound<- as.numeric(data1_tidy$Upper_bound)

#clean Maternal.Mortality
data1_tidy$Maternal.Mortality<-str_extract(data1_tidy$Maternal.Mortality, "[0-9]+\\[")
data1_tidy$Maternal.Mortality<-str_replace(data1_tidy$Maternal.Mortality, "\\[","")

data1_tidy$Year <- str_replace(data1_tidy$Year , "-[0-9]+" , "")

#change the type 
data1_tidy$Country<-as.character(data1_tidy$Country)
data1_tidy$Year<-as.numeric(data1_tidy$Year)
data1_tidy$Maternal.Mortality<-as.numeric(data1_tidy$Maternal.Mortality)
data1_tidy$Sklled.Health<-as.numeric(data1_tidy$Sklled.Health)

glimpse(data1_tidy)
```

- Download the gross national income data from http://apps.who.int/gho/data/node.main.GNI107?lang=en (Download the complete data as a "CVS table", storing it in the data folder).
- Tidy this data, so that there is one entry per country per year

```{r}
#use gather function
data2 <- read.csv("WHS9_93.csv" , skip = 1)

data2_tidy<-data2%>%
  gather(2:25 , key = "Year" , value = "Income")

data2_tidy$Year<-str_replace(data2_tidy$Year, "X.","")
data2_tidy$Income<-str_replace_all(data2_tidy$Income, " ", "")

#change the type
data2_tidy$Country<-as.character(data2_tidy$Country)
data2_tidy$Year<-as.numeric(data2_tidy$Year)
data2_tidy$Income<-as.numeric(data2_tidy$Income)

glimpse(data2_tidy)

```

- Merge the two files by country and year, retaining only rows for which you have per capita income and either maternal mortality or percentage attended births

```{r}
#Use merge to get the join data, then filter those rows !=NA.
data_join<-merge(data1_tidy , data2_tidy , by = c("Country"="Country" , "Year"="Year") ,all = TRUE)

data_join_tidy<-filter(data_join_tidy, Income !="NA")
data_join_tidy<-filter(data_join_tidy, Maternal.Mortality != "NA" |Sklled.Health != "NA")

glimpse(data_join_tidy)
```

- Make a two polished  and informative graphs, one relating income to maternal mortality and the other relating income to percentage attended births

```{r}
ggplot(data = data_join_tidy, aes(Maternal.Mortality , Income,color = Year )) + geom_point()

ggplot(data = data_join_tidy, aes(Sklled.Health, Income,color = Year )) + geom_point()
```

- Save a csv file with the merged, cleaned, and tidy data, using a suitable name of your choosing in the data folder.

```{r}
write.csv(data_join_tidy)
```

- Use the country_choropleth() function in the choroplethr package to make a world map of maternal mortality, using the most recent year for each country in the merged dataset you created. The defaults of the function will be fine; no need to tune up the plot. You can read the help file to see how the data must be formatted; you may need to rename the countries from the dataset that you've created.

```{r}
data_cho<-data_join_tidy%>%
  group_by(Country)%>%
  filter(Year== max(Year))

colnames(data_cho)[1] <- "region"
colnames(data_cho)[3] <- "value"

data_country <- subset(data_cho, select = c(region, value))
data_country$region <- tolower(data_country$region)

country_choropleth(data_country , "Maternal Mortality" )  

```

- All of your steps, other than downloading the files, should be contained in a single R Markdown file that will run in the repo directory. Be sure to describe your steps.